---
title: "Ratcalls_fda"
output: html_document
---
Entire analysis based on:http://www.statsoft.com/textbook/time-series-analysis
Methods: Only for rat calls: Spectrum Analysis
For other calls: Distributed Time Lags
Cheatsheet: Time series forecasting methods in Python: https://machinelearningmastery.com/time-series-forecasting-methods-in-python-cheat-sheet/

Maps: New York city census block - rats and other different sanitation calls
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Libraries
library(lubridate)
library(jsonlite)
library(plyr)
library(RJSONIO)
library(ggplot2)
library(plotly)
library(gridExtra)
library(reshape)
library(TTR)
library(forecast)
library(tseries)
library(vars)

options(stringsAsFactors = FALSE)
```
```{r}
raw_zip_monthly = fromJSON("/Users/kalyani/01ColumbiaQMSS/01Semester1/Practicum/data/ratcalls_zip_timeseries.json")
```

```{r}
#Format for access: [1][[1]][[4]][3]: count for fourth month-year of first zipcode 
#First [1] - zipcode
#Second [[1]]: keep, does not change (accesses elements)
#Third [[4]]: Accesses month-year
# Fourth [3]: Accesses count/month/year
#1: day, 2: month, 3: count
raw_zip_monthly[2][[1]][[4]][3]
```

```{r}
#Vector of counts:
raw_zip_counts = fromJSON("/Users/kalyani/01ColumbiaQMSS/01Semester1/Practicum/data/ratcalls_zip_timeseries_counts.json")
dat1 = as.data.frame(raw_zip_counts)
dat1$monthyear = c('(1, 2010)',
                   '(2, 2010)',
                   '(3, 2010)',
                   '(4, 2010)',
                   '(5, 2010)',
                   '(6, 2010)',
                   '(7, 2010)',
                   '(8, 2010)',
                   '(9, 2010)',
                   '(10, 2010)',
                   '(11, 2010)',
                   '(12, 2010)',
                   '(1, 2011)',
                   '(2, 2011)',
                   '(3, 2011)',
                   '(4, 2011)',
                   '(5, 2011)',
                   '(6, 2011)',
                   '(7, 2011)',
                   '(8, 2011)',
                   '(9, 2011)',
                   '(10, 2011)',
                   '(11, 2011)',
                   '(12, 2011)',
                   '(1, 2012)',
                   '(2, 2012)',
                   '(3, 2012)',
                   '(4, 2012)',
                   '(5, 2012)',
                   '(6, 2012)',
                   '(7, 2012)',
                   '(8, 2012)',
                   '(9, 2012)',
                   '(10, 2012)',
                   '(11, 2012)',
                   '(12, 2012)',
                   '(1, 2013)',
                   '(2, 2013)',
                   '(3, 2013)',
                   '(4, 2013)',
                   '(5, 2013)',
                   '(6, 2013)',
                   '(7, 2013)',
                   '(8, 2013)',
                   '(9, 2013)',
                   '(10, 2013)',
                   '(11, 2013)',
                   '(12, 2013)',
                   '(1, 2014)',
                   '(2, 2014)',
                   '(3, 2014)',
                   '(4, 2014)',
                   '(5, 2014)',
                   '(6, 2014)',
                   '(7, 2014)',
                   '(8, 2014)',
                   '(9, 2014)',
                   '(10, 2014)',
                   '(11, 2014)',
                   '(12, 2014)',
                   '(1, 2015)',
                   '(2, 2015)',
                   '(3, 2015)',
                   '(4, 2015)',
                   '(5, 2015)',
                   '(6, 2015)',
                   '(7, 2015)',
                   '(8, 2015)',
                   '(9, 2015)',
                   '(10, 2015)',
                   '(11, 2015)',
                   '(12, 2015)',
                   '(1, 2016)',
                   '(2, 2016)',
                   '(3, 2016)',
                   '(4, 2016)',
                   '(5, 2016)',
                   '(6, 2016)',
                   '(7, 2016)',
                   '(8, 2016)',
                   '(9, 2016)',
                   '(10, 2016)',
                   '(11, 2016)',
                   '(12, 2016)',
                   '(1, 2017)',
                   '(2, 2017)',
                   '(3, 2017)',
                   '(4, 2017)',
                   '(5, 2017)',
                   '(6, 2017)',
                   '(7, 2017)',
                   '(8, 2017)',
                   '(9, 2017)',
                   '(10, 2017)',
                   '(11, 2017)',
                   '(12, 2017)',
                   '(1, 2018)',
                   '(2, 2018)',
                   '(3, 2018)',
                   '(4, 2018)',
                   '(5, 2018)',
                   '(6, 2018)',
                   '(7, 2018)',
                   '(8, 2018)',
                   '(9, 2018)',
                   '(10, 2018)')
mdat1 = melt(dat1, id.vars="monthyear")
#Turn your 'treatment' column into a character vector
mdat1$monthyear <- as.character(mdat1$monthyear)
#Then turn it back into a factor with the levels in the correct order
mdat1$monthyear <- factor(mdat1$monthyear, levels=unique(mdat1$monthyear))
mdat1$variable = as.factor(mdat1$variable)
p = ggplot(mdat1, aes(x=monthyear, y=value)) +
    theme(panel.grid=element_blank()) +
    geom_line(aes(colour=variable),size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90)) + theme(legend.position="none")
ggsave("test.tiff", units="in", width=5, height=4, dpi=300, compression = 'lzw')
tiff("test.tiff", units="in", width=5, height=5, res=300)
dev.off()

```

```{r}
a <- list(
  type = "category",
  title = "",
  categoryorder = unique(mdat1$monthyear),
  showticklabels = FALSE,
  tickangle = 90,
  fontsize = 1
  )

b <- list(
  title = "Counts per Zipcode"
  )
p = plot_ly(mdat1, x = ~monthyear, y = ~value,color = ~variable, type = 'scatter',mode = 'lines', line = list(width = 0.75, alpha = 1)) %>% layout(xaxis = a, yaxis = b)
hide_legend(p)

```


```{r}
#Vector of counts:
raw_zip_counts = fromJSON("/Users/kalyani/01ColumbiaQMSS/01Semester1/Practicum/data/littercalls_zip_timeseries_counts.json")
dat2 = as.data.frame(raw_zip_counts)
dat2$monthyear = c('(1, 2010)',
       '(2, 2010)',
       '(3, 2010)',
       '(4, 2010)',
       '(5, 2010)',
       '(6, 2010)',
       '(7, 2010)',
       '(8, 2010)',
       '(9, 2010)',
       '(10, 2010)',
       '(11, 2010)',
       '(12, 2010)',
       '(1, 2011)',
       '(2, 2011)',
       '(3, 2011)',
       '(4, 2011)',
       '(5, 2011)',
       '(6, 2011)',
       '(7, 2011)',
       '(8, 2011)',
       '(9, 2011)',
       '(10, 2011)',
       '(11, 2011)',
       '(12, 2011)',
       '(1, 2012)',
       '(2, 2012)',
       '(3, 2012)',
       '(4, 2012)',
       '(5, 2012)',
       '(6, 2012)',
       '(7, 2012)',
       '(8, 2012)',
       '(9, 2012)',
       '(10, 2012)',
       '(11, 2012)',
       '(12, 2012)',
       '(1, 2013)',
       '(2, 2013)',
       '(3, 2013)',
       '(4, 2013)',
       '(5, 2013)',
       '(6, 2013)',
       '(7, 2013)',
       '(8, 2013)',
       '(9, 2013)',
       '(10, 2013)',
       '(11, 2013)',
       '(12, 2013)',
       '(1, 2014)',
       '(2, 2014)',
       '(3, 2014)',
       '(4, 2014)',
       '(5, 2014)',
       '(6, 2014)',
       '(7, 2014)',
       '(8, 2014)',
       '(9, 2014)',
       '(10, 2014)',
       '(11, 2014)',
       '(12, 2014)',
       '(1, 2015)',
       '(2, 2015)',
       '(3, 2015)',
       '(4, 2015)',
       '(5, 2015)',
       '(6, 2015)',
       '(7, 2015)',
       '(8, 2015)',
       '(9, 2015)',
       '(10, 2015)',
       '(11, 2015)',
       '(12, 2015)',
       '(1, 2016)',
       '(2, 2016)',
       '(3, 2016)',
       '(4, 2016)',
       '(5, 2016)',
       '(6, 2016)',
       '(7, 2016)',
       '(8, 2016)',
       '(9, 2016)',
       '(10, 2016)',
       '(11, 2016)',
       '(12, 2016)',
       '(1, 2017)',
       '(2, 2017)',
       '(3, 2017)',
       '(4, 2017)',
       '(5, 2017)',
       '(6, 2017)',
       '(7, 2017)',
       '(8, 2017)',
       '(9, 2017)',
       '(10, 2017)',
       '(11, 2017)',
       '(12, 2017)',
       '(1, 2018)',
       '(2, 2018)',
       '(3, 2018)',
       '(4, 2018)',
       '(5, 2018)',
       '(6, 2018)',
       '(7, 2018)',
       '(8, 2018)',
       '(9, 2018)',
       '(10, 2018)')
mdat2 = melt(dat2, id.vars="monthyear")

#Turn your 'treatment' column into a character vector
mdat2$monthyear <- as.character(mdat2$monthyear)
#Then turn it back into a factor with the levels in the correct order
mdat2$monthyear <- factor(mdat2$monthyear, levels=unique(mdat2$monthyear))
p = ggplot(mdat2, aes(x=monthyear, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_line(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))
```
```{r}
# Dirty Conditions Complaints:
#Vector of counts:
raw_zip_counts = fromJSON("/Users/kalyani/01ColumbiaQMSS/01Semester1/Practicum/data/dirtycalls_zip_timeseries_counts.json")
dat3 = as.data.frame(raw_zip_counts)
dat3$monthyear = c('(1, 2010)',
                 '(2, 2010)',
                 '(3, 2010)',
                 '(4, 2010)',
                 '(5, 2010)',
                 '(6, 2010)',
                 '(7, 2010)',
                 '(8, 2010)',
                 '(9, 2010)',
                 '(10, 2010)',
                 '(11, 2010)',
                 '(12, 2010)',
                 '(1, 2011)',
                 '(2, 2011)',
                 '(3, 2011)',
                 '(4, 2011)',
                 '(5, 2011)',
                 '(6, 2011)',
                 '(7, 2011)',
                 '(8, 2011)',
                 '(9, 2011)',
                 '(10, 2011)',
                 '(11, 2011)',
                 '(12, 2011)',
                 '(1, 2012)',
                 '(2, 2012)',
                 '(3, 2012)',
                 '(4, 2012)',
                 '(5, 2012)',
                 '(6, 2012)',
                 '(7, 2012)',
                 '(8, 2012)',
                 '(9, 2012)',
                 '(10, 2012)',
                 '(11, 2012)',
                 '(12, 2012)',
                 '(1, 2013)',
                 '(2, 2013)',
                 '(3, 2013)',
                 '(4, 2013)',
                 '(5, 2013)',
                 '(6, 2013)',
                 '(7, 2013)',
                 '(8, 2013)',
                 '(9, 2013)',
                 '(10, 2013)',
                 '(11, 2013)',
                 '(12, 2013)',
                 '(1, 2014)',
                 '(2, 2014)',
                 '(3, 2014)',
                 '(4, 2014)',
                 '(5, 2014)',
                 '(6, 2014)',
                 '(7, 2014)',
                 '(8, 2014)',
                 '(9, 2014)',
                 '(10, 2014)',
                 '(11, 2014)',
                 '(12, 2014)',
                 '(1, 2015)',
                 '(2, 2015)',
                 '(3, 2015)',
                 '(4, 2015)',
                 '(5, 2015)',
                 '(6, 2015)',
                 '(7, 2015)',
                 '(8, 2015)',
                 '(9, 2015)',
                 '(10, 2015)',
                 '(11, 2015)',
                 '(12, 2015)',
                 '(1, 2016)',
                 '(2, 2016)',
                 '(3, 2016)',
                 '(4, 2016)',
                 '(5, 2016)',
                 '(6, 2016)',
                 '(7, 2016)',
                 '(8, 2016)',
                 '(9, 2016)',
                 '(10, 2016)',
                 '(11, 2016)',
                 '(12, 2016)',
                 '(1, 2017)',
                 '(2, 2017)',
                 '(3, 2017)',
                 '(4, 2017)',
                 '(5, 2017)',
                 '(6, 2017)',
                 '(7, 2017)',
                 '(8, 2017)',
                 '(9, 2017)',
                 '(10, 2017)',
                 '(11, 2017)',
                 '(12, 2017)',
                 '(1, 2018)',
                 '(2, 2018)',
                 '(3, 2018)',
                 '(4, 2018)',
                 '(5, 2018)',
                 '(6, 2018)',
                 '(7, 2018)',
                 '(8, 2018)',
                 '(9, 2018)',
                 '(10, 2018)')

mdat3 = melt(dat3, id.vars="monthyear")
#Turn your 'treatment' column into a character vector
mdat3$monthyear <- as.character(mdat3$monthyear)
#Then turn it back into a factor with the levels in the correct order
mdat3$monthyear <- factor(mdat3$monthyear, levels=unique(mdat3$monthyear))
```
# Sanitary Complaints

```{r}
#Vector of counts:
raw_zip_counts = fromJSON("/Users/kalyani/01ColumbiaQMSS/01Semester1/Practicum/data/sanitarycalls_zip_timeseries_counts.json")
dat4 = as.data.frame(raw_zip_counts)
dat4$monthyear = c('(1, 2010)',
                   '(2, 2010)',
                   '(3, 2010)',
                   '(4, 2010)',
                   '(5, 2010)',
                   '(6, 2010)',
                   '(7, 2010)',
                   '(8, 2010)',
                   '(9, 2010)',
                   '(10, 2010)',
                   '(11, 2010)',
                   '(12, 2010)',
                   '(1, 2011)',
                   '(2, 2011)',
                   '(3, 2011)',
                   '(4, 2011)',
                   '(5, 2011)',
                   '(6, 2011)',
                   '(7, 2011)',
                   '(8, 2011)',
                   '(9, 2011)',
                   '(10, 2011)',
                   '(11, 2011)',
                   '(12, 2011)',
                   '(1, 2012)',
                   '(2, 2012)',
                   '(3, 2012)',
                   '(4, 2012)',
                   '(5, 2012)',
                   '(6, 2012)',
                   '(7, 2012)',
                   '(8, 2012)',
                   '(9, 2012)',
                   '(10, 2012)',
                   '(11, 2012)',
                   '(12, 2012)',
                   '(1, 2013)',
                   '(2, 2013)',
                   '(3, 2013)',
                   '(4, 2013)',
                   '(5, 2013)',
                   '(6, 2013)',
                   '(7, 2013)',
                   '(8, 2013)',
                   '(9, 2013)',
                   '(10, 2013)',
                   '(11, 2013)',
                   '(12, 2013)',
                   '(1, 2014)',
                   '(2, 2014)',
                   '(3, 2014)',
                   '(4, 2014)',
                   '(5, 2014)',
                   '(6, 2014)',
                   '(7, 2014)',
                   '(8, 2014)',
                   '(9, 2014)',
                   '(10, 2014)',
                   '(11, 2014)',
                   '(12, 2014)',
                   '(1, 2015)',
                   '(2, 2015)',
                   '(3, 2015)',
                   '(4, 2015)',
                   '(5, 2015)',
                   '(6, 2015)',
                   '(7, 2015)',
                   '(8, 2015)',
                   '(9, 2015)',
                   '(10, 2015)',
                   '(11, 2015)',
                   '(12, 2015)',
                   '(1, 2016)',
                   '(2, 2016)',
                   '(3, 2016)',
                   '(4, 2016)',
                   '(5, 2016)',
                   '(6, 2016)',
                   '(7, 2016)',
                   '(8, 2016)',
                   '(9, 2016)',
                   '(10, 2016)',
                   '(11, 2016)',
                   '(12, 2016)',
                   '(1, 2017)',
                   '(2, 2017)',
                   '(3, 2017)',
                   '(4, 2017)',
                   '(5, 2017)',
                   '(6, 2017)',
                   '(7, 2017)',
                   '(8, 2017)',
                   '(9, 2017)',
                   '(10, 2017)',
                   '(11, 2017)',
                   '(12, 2017)',
                   '(1, 2018)',
                   '(2, 2018)',
                   '(3, 2018)',
                   '(4, 2018)',
                   '(5, 2018)',
                   '(6, 2018)',
                   '(7, 2018)',
                   '(8, 2018)',
                   '(9, 2018)',
                   '(10, 2018)')
mdat4 = melt(dat4, id.vars="monthyear")
#Turn your 'treatment' column into a character vector
mdat4$monthyear <- as.character(mdat4$monthyear)
#Then turn it back into a factor with the levels in the correct order
mdat4$monthyear <- factor(mdat4$monthyear, levels=unique(mdat4$monthyear))
p = ggplot(mdat4, aes(x=monthyear, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_line(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))
```



## Calculating variance/range by zipcode - boxplots
## Scatter plot of means with variance


# Simple moving average
# For one zip code

```{r}
mdat1_X10027 = subset(mdat1, variable == "X10027")
mdat1_X11201$sma = SMA(mdat1_X11201$value)
mdat1_X11201$sma
mdat1_X11201 = mdat1_X11201[-(0:9), , drop = FALSE]
ggplot(mdat1_X11201, aes(monthyear, group = 1)) + geom_line(aes(y=value, colour = "var0")) + geom_line(aes(y=sma, colour = "var1")) 
```



# Exponential moving average
# For one zip code
```{r}
mdat1_X11201$ema = EMA(mdat1_X11201$value)
mdat1_X11201 = mdat1_X11201[-(0:9), , drop = FALSE]
ggplot(mdat1_X11201, aes(monthyear, group = 1)) + geom_line(aes(y=value, colour = "var0")) + geom_line(aes(y=ema, colour = "var1")) 
```


https://www.rdocumentation.org/packages/forecast/versions/8.4/topics/Acf


# Cross Correlation - can be performed

```{r}
acf = Acf(mdat1_X10027$value, plot = TRUE)
acf = unlist(acf, recursive = TRUE, use.names = FALSE)
acf = acf[0:13]
acf
```

#  ACFs - Scatter-plot
```{r}
dat1 <- dat
i <-  0

mat1 = matrix(0,37,1)
dim(mat1)
for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    acf <- Acf(dat1[column], lag.max = 36)
    acf = unlist(acf, recursive = TRUE, use.names = FALSE)
    acf = acf[0:37]
    mat1 = cbind(mat1, acf)
  i = i + 1
  }
}

df_acf = data.frame(mat1)
drops <- c("V1")
df_acf = df_acf[ , !(names(df_acf) %in% drops)]

dat2 = dat[,1:203]
head(dat2)
df_acf = df_acf[,1:203]
colnames(df_acf) = colnames(dat2)
df_acf$lag = seq(0,36,1)

df_acf = melt(df_acf, id.vars="lag")
df_acf
df_acf$value = as.numeric(df_acf$value)

p = ggplot(df_acf, aes(x=lag, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_point(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))

a <- list(
  type = "category",
  title = "lag",
  showticklabels = TRUE,
  tickangle = 90,
  fontsize = 1
  )
  
p = plot_ly(df_acf, x = ~lag, y = ~value,color = ~variable, type = 'scatter',marker = list(size = 3, alpha = 0.5, line = list(width = 2))) %>% layout(xaxis = a)
p
```


```{r}
pacf = Pacf(mdat1_X10027$value, plot = TRUE)
```

# Differencing to remove serial dependency for exponential smoothing
```{r}
value = mdat1_X10027$value

one_diff <- function(values){
  x <- c(1,2)
  for (i in 1:length(values)) {
    v = value[i] - value[i-1]
    x <- append(x,v)
    i = i+1
  }
  return(x)
}

value_lag1 = one_diff(value)
value_lag1 = value_lag1[-(0:2)]
mdat1_X10027 = mdat1_X10027[-c(0:1), ]
mdat1_X10027['value_lag1'] = value_lag1

ggplot(mdat1_X10027, aes(monthyear, group = 1)) + geom_line(aes(y=value, colour = "red")) + geom_line(aes(y=value_lag1, colour = "blue")) 
acf = Acf(mdat1_X10027$value_lag1, plot = TRUE)
```

#  PACFs - Scatter-plot
```{r}
dat1 <- dat
i <-  0

mat1 = matrix(0,37,1)
dim(mat1)
for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    acf <- Pacf(dat1[column], lag.max = 36)
    acf = unlist(acf, recursive = TRUE, use.names = FALSE)
    acf = acf[0:37]
    mat1 = cbind(mat1, acf)
  i = i + 1
  }
}

df_acf = data.frame(mat1)
drops <- c("V1")
df_acf = df_acf[ , !(names(df_acf) %in% drops)]

dat2 = dat[,1:203]
head(dat2)
df_acf = df_acf[,1:203]
colnames(df_acf) = colnames(dat2)
df_acf$lag = seq(0,36,1)

df_acf = melt(df_acf, id.vars="lag")
df_acf
df_acf$value = as.numeric(df_acf$value)

p = ggplot(df_acf, aes(x=lag, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_point(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))

plot_ly(df_acf, x = ~lag, y = ~value,color = ~variable, type = 'scatter',marker = list(size = 3, alpha = 0.5, line = list(width = 2))) %>% layout(xaxis = a)
```
## Exponential smoothing for rat calls (first differencing)
```{r}
dat1 <- dat
mat1 = matrix(0,105,1)
cols <- c(0)
length(cols)
dim(mat1)
i <-  2
one_diff <- function(df, column){
  x <- c(1,2)
  while (i < 106) {
    v = df[column][i, ] - df[column][i-1, ]
    x = append(x,v)
    i = i+1
  }
  return(x)
}

j <-  0

for (column in colnames(dat1)) {
  if (column != "monthyear") {
    if (!(column %in% cols)) {
    cols[j+1] = column
    value_lag1 = one_diff(dat1, column)
    value_lag1 = value_lag1[-(0:2)]
    mat1 = cbind(mat1, value_lag1)
  j = j + 1
    }
  }
}
dim(mat1)
length(cols)

df_lag1 = data.frame(mat1)

drops <- c("V1")
df_lag1 = df_lag1[ , !(names(df_lag1) %in% drops)]
colnames(df_lag1) = cols
df_lag1$time = seq(2, 106, by=1)
df_lag1 = melt(df_lag1, id.vars="time")
df_lag1$value = as.numeric(df_lag1$value)
df_lag1
df_lag1$variable = as.factor(df_lag1$variable)

p = ggplot(df_lag1, aes(x=time, y=value, color=variable)) +
    theme(panel.grid=element_blank()) +
    geom_line(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90)) + theme(legend.position="none")
p

a <- list(
  type = "category",
  categoryorder = unique(df_lag1$time),
  showticklabels = TRUE,
  tickangle = 90,
  fontsize = 3
  )

b <- list(range = c(-70, 70))

plot_ly(df_lag1, x = ~time, y = ~value,color = ~variable, type = 'scatter',mode = 'lines', line = list(width = 0.75, alpha = 1)) %>% layout(xaxis = a, yaxis = b)

s = ses(df_lag1_10027[0:75,3], h=20,  alpha=0.1, initial="simple")
accuracy(s, df_lag1_10027[76:95,3] )

```


```{r}
df_lag1_1 = data.frame(mat1)
drops <- c("V1")
df_lag1_1 = df_lag1_1[ , !(names(df_lag1_1) %in% drops)]
dim(df_lag1_1)
df_lag1_1
forecasts = c()
for (column in colnames(df_lag1_1)) {
  if (column != "monthyear") {
    if (!(column %in% cols)) {
    s = ses(df_lag1_1[column], h=12,  alpha=0.1, initial="simple")
    num = round(s['mean'][[1]][[1]], 2)
    forecasts = append(forecasts, num)
    }
  }
}
length(forecasts)
mat2 = cbind(cols, forecasts)
df_ses = as.data.frame(mat2, stringsAsFactors = FALSE)
df_ses$cols <- factor(df_ses$cols, levels = unique(df_ses$cols)[order(df_ses$forecasts, decreasing = TRUE)])
df_ses
plot_ly(df_ses, x = ~cols, y = ~forecasts, type = 'bar') %>%
  layout(yaxis = list(title = 'Forecast'))


```



# CCFs with litter baskets
```{r}
dat1 <- dat
i <-  0
mat1 = matrix(0,73,1)
cols <- c(0)
length(cols)
dim(mat1)
for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    if (column %in% colnames(dat2)) {
    if (!(column %in% cols)) {
    cols[i+1] = column
    ccf <- Ccf(dat1[column], dat2[column], lag.max = 36)
    ccf = unlist(ccf, recursive = TRUE, use.names = FALSE)
    ccf = ccf[0:73]
    mat1 = cbind(mat1, ccf)
  i = i + 1
    }
    }
  }
}


df_ccf = data.frame(mat1)
drops <- c("V1")
df_ccf = df_ccf[ , !(names(df_ccf) %in% drops)]
colnames(df_ccf) = cols
df_ccf$lag = seq(-36, 36, by=1)

df_ccf = melt(df_ccf, id.vars="lag")
df_ccf$value = as.numeric(df_ccf$value)

p = ggplot(df_ccf, aes(x=lag, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_point(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))

plot_ly(df_ccf, x = ~lag, y = ~value,color = ~variable, type = 'scatter',marker = list(size = 3, alpha = 0.5, line = list(width = 2))) %>% layout(xaxis = a)


```
# CCFs with dirty conditions
```{r}
dat1 <- dat
i <-  0
mat1 = matrix(0,73,1)
cols <- c(0)
length(cols)
dim(mat1)
for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    if (column %in% colnames(dat3)) {
    if (!(column %in% cols)) {
    cols[i+1] = column
    ccf <- Ccf(dat1[column], dat3[column], lag.max = 36)
    ccf = unlist(ccf, recursive = TRUE, use.names = FALSE)
    ccf = ccf[0:73]
    mat1 = cbind(mat1, ccf)
  i = i + 1
    }
    }
  }
}

cols
length(cols)
cols
df_ccf = data.frame(mat1)
dim(df_ccf)
drops <- c("V1")
df_ccf = df_ccf[ , !(names(df_ccf) %in% drops)]
dim(df_ccf)
head(df_ccf)
colnames(df_ccf) = cols
df_ccf$lag = seq(-36, 36, by=1)

df_ccf = melt(df_ccf, id.vars="lag")
df_ccf
df_ccf$value = as.numeric(df_ccf$value)

p = ggplot(df_ccf, aes(x=lag, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_point(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))

plot_ly(df_ccf, x = ~lag, y = ~value,color = ~variable, type = 'scatter',marker = list(size = 3, alpha = 0.5, line = list(width = 2))) %>% layout(xaxis = a)


```

# CCFs with sanitary conditions
```{r}
dat1 <- dat
i <-  0
mat1 = matrix(0,73,1)
cols <- c(0)
length(cols)
dim(mat1)
for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    if (column %in% colnames(dat4)) {
    if (!(column %in% cols)) {
    cols[i+1] = column
    ccf <- Ccf(dat1[column], dat4[column], lag.max = 36)
    ccf = unlist(ccf, recursive = TRUE, use.names = FALSE)
    ccf = ccf[0:73]
    mat1 = cbind(mat1, ccf)
  i = i + 1
    }
    }
  }
}

cols
length(cols)
cols
df_ccf = data.frame(mat1)
dim(df_ccf)
drops <- c("V1")
df_ccf = df_ccf[ , !(names(df_ccf) %in% drops)]
dim(df_ccf)
head(df_ccf)
colnames(df_ccf) = cols
df_ccf$lag = seq(-36, 36, by=1)

df_ccf = melt(df_ccf, id.vars="lag")
df_ccf
df_ccf$value = as.numeric(df_ccf$value)

p = ggplot(df_ccf, aes(x=lag, y=value, group=variable)) +
    theme_bw() +
    theme(panel.grid=element_blank()) +
    geom_point(size=0.5, alpha=0.3) + theme(text = element_text(size=6), axis.text.x = element_text(angle = 90))

plot_ly(df_ccf, x = ~lag, y = ~value,color = ~variable, type = 'scatter',marker = list(size = 3, alpha = 0.5, line = list(width = 2))) %>% layout(xaxis = a)
```

# Var model testing

```{r}
mat1 = cbind(dat1['X10027'], dat2['X10027'], dat3['X10027'], dat4['X10027'])
dim(mat1)
colnames(mat1) = c('rats', 'litter', 'dirty', 'sanitary')
var <- VAR(mat1, 12)
summary(var)
library(corrplot)
source("http://www.sthda.com/upload/rquery_cormat.r")
rquery.cormat(mat1, type = "full")



predict(var, n.ahead = 1)[[1]][[1]][[1]] #forecast
predict(var, n.ahead = 1)[[1]][[1]][[2]] #lower
predict(var, n.ahead = 1)[[1]][[1]][[3]] #higher
predict(var, n.ahead = 1)[[1]][[1]][[4]] #CI

# R squared
summary(var)[2][[1]][[1]][[9]]
summary(var)[2][[1]][[2]][[9]]
summary(var)[2][[1]][[3]][[9]]
summary(var)[2][[1]][[4]][[9]]
```

# Var Model

```{r}
i <-  0
cols <- c()
length(cols)
dim(mat1)

pred_rats = c()
pred_rats_CI = c()

pred_litter = c()
pred_litter_CI = c()

pred_dirty = c()
pred_dirty_CI = c()

pred_sanitary = c()
pred_sanitary_CI = c()

for (column in colnames(dat1)) {
  c = as.character(i)
  if (column != "monthyear") {
    if (column %in% colnames(dat2)) {
      if (column %in% colnames(dat3)) {
        if (column %in% colnames(dat4)) {

    if (!(column %in% cols)) {
    cols[i+1] = column
    mat1 = cbind(dat1[column], dat2[column], dat3[column], dat4[column])
    colnames(mat1) = c('rats', 'litter', 'dirty', 'sanitary')
    var <- VAR(mat1, lag.max = 12)
    pred = predict(var, n.ahead = 1)
    pred_rats = append(pred_rats, pred[[1]][[1]][[1]])
    pred_rats_CI = append(pred_rats_CI, pred[[1]][[1]][[4]])
    
    pred_litter = append(pred_rats, pred[[1]][[2]][[1]])
    pred_litter_CI = append(pred_rats_CI, pred[[1]][[2]][[4]])
    
    pred_dirty = append(pred_dirty, pred[[1]][[3]][[1]])
    pred_dirty_CI = append(pred_rats_CI, pred[[1]][[3]][[4]])
    
    pred_sanitary = append(pred_sanitary, pred[[1]][[4]][[1]])
    pred_sanitary_CI = append(pred_sanitary_CI, pred[[1]][[4]][[4]])
  i = i + 1
    }
        }
      }
    }
  }
}

mat_preds_var = cbind(cols, pred_rats, pred_rats_CI, pred_litter, pred_litter_CI, pred_dirty, pred_dirty_CI, pred_sanitary, pred_sanitary_CI)

df_preds_var = as.data.frame((mat_preds_var))
df_preds_var
```

